---
title: "Practical Machine Learning Project"
author: "Peiran Chen"
date: "17/06/2017"
output: 
  html_document: 
    keep_md: yes
---

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

## Executive Summary
- Load training and testing data
- Subdivide training data for cross validation
- Train data using different model
- Predict on test data

## Load training and testing data
```{r}
# load required libraries
library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
# set seed for reproducibility
set.seed(1234)
# load training data and set NAs for missing values
training <- read.csv("/Users/mq20120117/Documents/git-repo/practicalmachinelearning/data/pml-training.csv", na.strings=c("NA","#DIV/0!", ""))
# load testing data and set NAs for missing values
testing <- read.csv("/Users/mq20120117/Documents/git-repo/practicalmachinelearning/data/pml-testing.csv" , na.strings=c("NA","#DIV/0!", ""))

# check str(training) and find many variables are NA. Remove them from data set
# Some variables are not related with our model, such as X, user_name, raw_timestamp_part_1,
# raw_timestamp_part_2, cvtd_timestamp, cvtd_timestamp, new_window and num_window, column 1-7
training <- training[,colSums(is.na(training)) == 0]
training <-training[,-c(1:7)]
testing <- testing[,colSums(is.na(testing)) == 0]
testing <- testing[,-c(1:7)]

# Take a look at our data
dim(training)
head(training)
dim(testing)
head(testing)
```

## Subdivide training data for cross validation

We have 19,622 observations in our training data set. And 20 observations in test data set.
Subdivide training data set into sub-training and sub-testing.
It will be random sampling without replacement
```{r}
sub.index <- createDataPartition(y=training$classe, p=0.75, list=FALSE)
sub.training <- training[sub.index, ] 
sub.testing <- training[-sub.index, ]
dim(sub.training)
dim(sub.testing)
```

## Train data using different model

Check frequency of outcomes
```{r}
plot(sub.training$classe)
```

From the above chart, we know each category have the similar size in our testing data set.

### Random forest model

```{r}
random.forest.model <- randomForest(classe ~. , data=sub.training, method="class")
random.forest.predict <- predict(random.forest.model, sub.testing, type = "class")
# test result
confusionMatrix(random.forest.predict, sub.testing$classe)
```

### Decision tree model

```{r}
decision.tree.model <- rpart(classe ~ ., data=sub.training, method="class")
decision.tree.predict <- predict(decision.tree.model, sub.testing, type = "class")
# plot tree
rpart.plot(decision.tree.model)
# test result
confusionMatrix(decision.tree.predict, sub.testing$classe)
```

By comparison, random forest model accuracy is 0.9953 vs. decision tree model accuracy is 0.7394. So we will choose random forest model.

## Predict on test data

We use the random forest model to predict our test data set.
```{r}
test.predict <- predict(random.forest.model, testing, type="class")
test.predict
```